A Large-scale Study of Polymorphism in R

### 2019-03-21 paper/main.tex
implementation

@@

The contributions of this paper are thus as follows:
\begin{itemize}
\item A large-scale analysis of the polymorphism present in function
  signatures of \PACKAGES widely used and actively maintained R packages.
\item A tracing and analysis pipeline that extends a previously published
  test generation tool named \code{genthat}.
\item Manual analysis of 100 functions to validate the dynamic analysis
  results.
\end{itemize}

@@

\newcommand{\genthat}{{\sc genthat}\xspace}

@@

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
`+`: D[1], I[1] -> D[1]
`c`: D[1], D[1], D[1], D[1] -> D[4]
`<-`: S, D[3] -> D[3]
`c`: D[1], D[1] -> D[2]
`attr<-1`:  S, C, D[2] -> 
`c`: D[1], D[1] -> D
`+`: D[4]{dim=(2,2)}, D[2] -> D[4]{dim=(2,2)}
\end{lstlisting}}\caption{Recorded shapes}\label{shapes}\end{figure}

@@

>>>>

The high-level process for tracing one package is as follows:

\begin{enumerate}
\item first, all functions in the package are instrumented with on-exit
  hooks.  These hooks call a number of R's available reflection functions on
  each of the arguments, as well as the return type: The called functions
  are {\tt typeof}, {\tt class}, and {\tt attributes}.  Recording is done on
  function exit to avoid forcing promises before they are needed (as these
  promises may cause side effects).

\item next, runnable code is extracted and run from package tests, examples,
  and vignettes.

\item each function invocation triggers the on-exit hook, and it is this
  on-exit code which collects the information.  Depending on the results of
  {\tt typeof}, {\tt class}, and {\tt attributes}, more analysis may be
  performed.  For example, if an argument is a list, type information is
  collected on list elements until a more accurate type (e.g., {\tt
    list<double>}) can be constructed.  Indeed, in a later section we do
  propose some possible type systems for R, and our analysis can be
  parameterized with these type systems to generate type information
  according to the types specified, and this will be discussed further in
  Section \AT{yap}.

\item all of the recorded type information makes up a function {\it trace}.
  This trace is passed to a C++ function (for performance reasons) which
  hashes it and checks it against already seen trace hashes.  If the trace
  is determined to be new, the hash is saved and the trace is written to a
  file.

\item finally, after all test code has finished executing, we are left with
  a directory full of all of the unique type signatures that were observed
  while running the package code.
\end{enumerate}

For each package function, we then consolidate all observed signatures: this
gives us a {\it function signature}.  \AT{Example?}  In the function
signature, each function argument is accompanied by a list of types,
attributes, and classes that have inhabited it.

>>>>

\newcommand{\genthat}{{\sc Genthat}\xspace}

@@

The contributions of this paper are thus as follows:
\begin{itemize}
\item A large-scale analysis of the polymorphism present in function
  signatures of \PACKAGES widely used and actively maintained R packages.
\item A tracing and analysis pipeline that extends a previously published
  test generation tool named \genthat.
\item Manual analysis of 100 functions to validate the dynamic analysis
  results.
\end{itemize}

@@

@@

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
`+`: D[1], I[1] -> D[1]
c: D[1], D[1], D[1], D[1] -> D[4]
`<-`: S, D[3] -> D[3]
c: D[1], D[1] -> D[2]
`attr<-1`:  S, C, D[2] -> 
c: D[1], D[1] -> D[2]
`+`: D[4]{dim=D[2]}, D[2] -> D[4]{dim=D[2]}
\end{lstlisting}}\caption{Recorded shapes}\label{shapes}\end{figure}

@@

A high-level description of the workflow of our tool for one package
retrieved from CRAN is as follows:

\begin{enumerate}
\item {\bf Exec generation:} All runnable code in the package is extracted
  from its tests, examples and vignettes. The code snippets are combined
  into a single file.
\item {\bf Installation:} All packages that are required for execution of
  the current package are downloaded and installed.
\item {\bf Instrumentation:} As code is loaded into the R, every function
  definition is instrumented with an \code{on-exit} hook which is invoked
  when the function returns either normally or through an exception.
\item {\bf Recording:} When a hook is called, arguments and return value are
  inspected. We record \k{typeof}, \k{class} and \k{attributes} recursively.
  For \code{list} values, an extra bit of analysis is performed to record
  the element type.
\item {\bf Writing:} Unique signatures are recorded to file with information
  about which package triggered the recording.
\end{enumerate}

Our recording mechanism does not remember the order in which arguments were
passed, nor does it record which arguments were not passed (and for which
default values were used). If an argument was not passed, and no default
value was specified, we record it as \emph{missing}. In R, trying to use
such a missing value results in an error. For practical reasons, we do not
record the contents of environments. These can be used as hash tables and
may be big and are quite likely different from one another.

In R all arguments are passed as promises, unused arguments will be
unevaluated, for these we can not report a type. THUS WE DO WHAT? SAY MORE.

For each function of each package, we consolidate all recorded observations
into a function signature. The details of how consolidation is parameterized
by the definition of a potential type system for R and will be discussed in
the next section.

\section{Stuff}
